---
title: "Ficus_Sycamorus_Experiment"
author: "Emmanuel Lourie"
date: "2024-01-25"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls()) # clean history 
library(tidyverse)
```

Get data of tree-visits (routine movement) and of the manipulation results (documented manually). 
```{r}
all_stops<-read_csv("data/all_tree_visits.csv") # Table of all tree visits ("stops") 
all_stops<-all_stops[all_stops$dummy!="cave",] # Select only stops on trees, without cave-roosts (dummy=cave)

syc_stops_north<-read_csv("data/visits_to_f.sycamorus.csv") # Tree visits within 100 m buffer around the Northern target F.sycomorus tree (choosing stops within this buffer was done in ArcGIS)

syc_stops_south<-all_stops[all_stops$cluster_ID=="Ficus sycomorus_2",]# Because the southern tree is in close proximity with other relevant fruit trees, the delineation of a visit to it is based on a harder criterion, which is based on a previous analysis of the track segmentation and then binning of the relative number of points (X, Y coordinates that were classified as a 1-tree stop) that are closest to this tree more than other close-by trees. 
cols<-colnames(syc_stops_north)
syc_stops_south<-syc_stops_south[,cols]
syc_stops<-rbind(syc_stops_south,syc_stops_north) # All tree stops to one of the two F. sycomorus trees during "routine" foraging movements. 
```

Get the table with the field manipulation results.
The table includes visits to the two F. sycomorus, that the bats exposed to the field manipulation made. Counting visits was made manually from observing the tracks 0-3 nights after the manipulation. 
```{r}
manip<-read_csv("data/Field_Manipulation_Table.csv")
```

Comparisons to visits to F. sycomorus during routine movements, for periods of known phenology and not during the manipulation. Here, I create a dataset that merges all tag-nights that visited the F. sycomorus trees (syc_stops_unq) and the ones that did not (stops_no_syc). i.e. the rest of the tag-nights. 
```{r}

syc_tag_nights<-unique(syc_stops$TAG_Night)
stops_no_syc<-all_stops[!all_stops$TAG_Night %in% syc_tag_nights,]
stops_no_syc<-stops_no_syc[!duplicated(stops_no_syc$TAG_Night),]
stops_no_syc$syc<-rep("no",nrow(stops_no_syc))
syc_stops_unq<-syc_stops[!duplicated(syc_stops$TAG_Night),]
syc_stops_unq$syc<-rep("yes",nrow(syc_stops_unq))
cols<-colnames(syc_stops_unq)
stops_no_syc<-stops_no_syc[,cols]

syc_dt<-rbind(syc_stops_unq, stops_no_syc)
table(syc_dt$syc)
```

Tables for dates of different phenological stages of the F. sycomorus trees (based on field monitoring).
```{r}
phen_table <- read.table(text = "  Start        End  stage
1	2019-11-11	2020-01-14	no_fruits
2	2020-01-15	2020-06-01	NA
3	2020-06-02	2020-06-12	fruits
4	2020-06-13	2020-06-22	no_fruits
5	2020-06-23	2020-06-25	manipulation
6	2020-06-26	2020-06-30	no_fruits
7	2020-07-01	2020-11-01	fruits
8	2020-11-02	2020-11-30	NA
9	2020-12-01	2021-03-30	no_fruits
11	2019-12-15	2020-02-15	no_fruits
12	2018-12-15	2019-02-15	no_fruits", # Add winters artificially
header = TRUE,
stringsAsFactors = FALSE)

phen_table$Start<-as.Date(phen_table$Start, format="%Y-%m-%d")
phen_table$End<-as.Date(phen_table$End, format="%Y-%m-%d")

phen_start<-min(phen_table$Start)
phen_end<-max(phen_table$End)


```


# During the manipulation, some bats racted immediatly by arriving to the target tree, while some visited it up to 3 nights after. Hence, I asked whther there was a visit within 3 nights interval (and later compare the probability of visiting the target tree witihn 3-nights interval during non-manipulation periods)
```{r}
syc_dt<-syc_dt[!is.na(syc_dt$date_global),]
start_date<-min(syc_dt$date_global); end_date<-max(syc_dt$date_global)

syc_dt_all<-syc_dt
man_start<-na.omit(phen_table$Start[phen_table$stage=="manipulation"])
man_end<-na.omit(phen_table$End[phen_table$stage=="manipulation"])
before_man<-seq(man_start-1,start_date, -3)
after_man<-seq(man_end+1, end_date,3)
man<-seq(man_start,man_end,3)
triple_dates<-c(before_man,man,after_man)
triple_inx<-seq(1, length(triple_dates),1)
triple_df<-data.frame("start_date"=triple_dates, "triple_inx"=triple_inx)

# add penology
syc_dt_all$fruits<-NA
for (r in 1:nrow(phen_table)){
rows<-which(syc_dt_all$date_global>=phen_table$Start[r] & syc_dt_all$date_global<=phen_table$End[r])
syc_dt_all$fruits[rows]<-phen_table$stage[r]
}
table(syc_dt_all$fruits, useNA = "ifany")

triple_df<-triple_df[order(triple_df$start_date),]
tdate_l<-unique(triple_df$start_date)
syc_dt_all$triple_date<-rep(NA,nrow(syc_dt_all))

for (t in 1:length(tdate_l)){
triple_loc<-which(syc_dt_all$date_global>= tdate_l[t] & syc_dt_all$date_global< tdate_l[t+1])
if(length(triple_loc)>0){
syc_dt_all$triple_date[triple_loc]<-triple_df$triple_inx[triple_df$start_date==tdate_l[t]]
 }
}

syc_dt_all$TAG_triple<-paste0(syc_dt_all$TAG,"_", syc_dt_all$triple_date)
syc_dt_all<-syc_dt_all %>% group_by(triple_date) %>% mutate("tdate"=min(as.Date(date_global)))

# Remove manipulation dates:

syc_dt_all<-syc_dt_all[syc_dt_all$fruits!="manipulation",]

```

Summarize - what is the proportion of visits to F. sycomorus (tags visiting the tree relative to all tags).Remove the manipulation period.  
```{r}
syc_dt_sum_t<-syc_dt_all %>% group_by(fruits,TAG, tdate) %>% summarise("n_syc"=length(syc[syc=="yes"]))
syc_dt_sum_t$syc_vis<-ifelse(syc_dt_sum_t$n_syc>0,1,0)
table(syc_dt_sum_t$fruits)

syc_sum_triple<-syc_dt_sum_t %>% group_by(fruits,tdate) %>% summarise("n_syc_vis"=sum(syc_vis), "n_tags"=length(unique(TAG)))
syc_sum_triple$syc_prop<-syc_sum_triple$n_syc_vis/syc_sum_triple$n_tags
syc_sum_triple<-syc_sum_triple[!is.na(syc_sum_triple$fruits),] # removing dates without phenological monitoring


```

Chi-square test - comparing the probability to visit the target trees when there are no fruits to the probability during the manipulation (when, also, there were no fruits on the trees).

```{r}

syc_sum_test<-syc_sum_triple %>% group_by(fruits) %>% summarise("n_vis" = sum(n_syc_vis),"n_tags"=sum(n_tags))
# Use only times when there were no fruits:
syc_sum_test_nof<-syc_sum_test[syc_sum_test$fruits=="no_fruits",]
# Add manipulation results
# Note that one date of the manipulation was made in Gershom cave, but not in Zemer, so there is no data for it (NA)/

syc_man<-data.frame("fruits"="manipulation" ,"n_vis"=sum(manip$n_visited_either_tree,na.rm=T)  ,"n_tags"=sum(manip$n_total,na.rm=T))



syc_nof_test<-rbind(syc_man,syc_sum_test_nof)
contingency_table_nof <- xtabs(cbind(n_vis, n_tags) ~ fruits, data = syc_nof_test)
print(chisq.test(contingency_table_nof))

print(paste0("proportion of mainpulated bats visiting one of the trees = ",round(syc_man$n_vis/syc_man$n_tags,2)))
print(paste0("mean proportion of bats visiting one of the trees when there are no fruits= ",round(syc_sum_test_nof$n_vis/syc_sum_test_nof$n_tags,2)))

```
 
