---
title: "Ficus Sycamorus Experiment"
author: "Emmanuel Lourie" 
output: html_document
date: "2023-09-19"
keep_md: true
chunk_output_type: console
editor_options: 
  chunk_output_type: console
---

**Clean history and load libraries**
```{r setup, include=FALSE}
rm(list=ls()) # clean history 
library(tidyverse)
options(dplyr.summarise.inform = FALSE)
```

**Get data of tree-visits to the focal trees (all periods)** 
```{r}
all_stops<-read_csv("data_R/all_tree_visits_main.csv") # Table of all tree visits ("stops") 
all_stops<-all_stops[all_stops$dummy!="cave",] # Filter out the cave-roost visits from the dataset to focus on tree visitation patterns (dummy=cave)ץ

syc_stops_north<-read_csv("data_R/bats_visits_north_ficus_30_m_Radius.csv") # Tree visits within 30 m buffer around the Northern target F.sycomorus tree 

syc_stops_south<-all_stops[all_stops$cluster_ID=="Ficus sycomorus_2",]# Because the southern tree is in close proximity with other relevant fruit trees, the delineation of a visit to it is based on a harder criterion, which is based on a previous analysis of the track segmentation and then binning of the relative number of points (X, Y coordinates that were classified as a 1-tree stop) that are closest to this tree more than other close-by trees. 

cols<-colnames(syc_stops_south)
syc_stops_north<-syc_stops_north[,cols]
syc_stops<-rbind(syc_stops_south,syc_stops_north) # All tree stops to one of the two F. sycomorus trees during "routine" foraging movements. 
```

**Get the table with the field manipulation results.**
**The table includes visits to the two *F. sycomorus*, that the bats exposed to the field manipulation made. Counting visits was made manually from observing the tracks 0-3 nights after the manipulation.** 
```{r}
manip<-read_csv("data_R/Field_Manipulation_Table.csv")
manip_syc_tags<-read_csv("data_R/Tags_manipulated_and_visiting.csv")
tag_smeared<-c("6485", "6637","6654","6639")
manip_syc_tags$manip<-ifelse(manip_syc_tags$tags%in% tag_smeared, "Manipulated","Naive")
```

**Comparisons to visits to *F. sycomorus* during routine movements, for periods of no fruits and not during the manipulation. Here, I create a dataset that merges all tag-nights that visited the *F. sycomorus* trees (syc_stops_unq) and the ones that did not (stops_no_syc). *i.e.* the rest of the tag-nights.** 
```{r}

syc_tag_nights<-unique(syc_stops$TAG_Night)
stops_no_syc<-all_stops[!all_stops$TAG_Night %in% syc_tag_nights,]
stops_no_syc<-stops_no_syc[!duplicated(stops_no_syc$TAG_Night),]
stops_no_syc$syc<-rep("no",nrow(stops_no_syc))
syc_stops_unq<-syc_stops[!duplicated(syc_stops$TAG_Night),]
syc_stops_unq$syc<-rep("yes",nrow(syc_stops_unq))
cols<-colnames(syc_stops_unq)
stops_no_syc<-stops_no_syc[,cols]

syc_dt<-rbind(syc_stops_unq, stops_no_syc)
# Classify date_global as date format
syc_dt$date_global<-as.Date(syc_dt$date_global, format = "%d/%m/%Y")
str(syc_dt)
table(syc_dt$syc)
# the date is "global" to give each night a single date from sunset to sunrise. 
```

**Tables for dates of different phenological stages of the *F. sycomorus* trees (based on field monitoring).**
```{r}
phen_table <- read.table(text = "  Start        End  stage
1	2019-11-11	2020-01-14	no_fruits
2	2020-01-15	2020-06-01	NA
3	2020-06-02	2020-06-12	fruits
4	2020-06-13	2020-06-22	no_fruits
5	2020-06-23	2020-06-25	manipulation
6	2020-07-01	2020-11-01	fruits
7	2020-11-02	2020-11-30	NA
8	2020-12-01	2021-03-30	no_fruits
9	2019-12-15	2020-02-15	no_fruits
10	2018-12-15	2019-02-15	no_fruits", # Add winter
header = TRUE,
stringsAsFactors = FALSE)

phen_table$Start<-as.Date(phen_table$Start, format="%Y-%m-%d")
phen_table$End<-as.Date(phen_table$End, format="%Y-%m-%d")

phen_table$n_nights<-phen_table$End-phen_table$Start

phen_start<-min(phen_table$Start)
phen_end<-max(phen_table$End)


```


###Comparing the manipulation results to the routine movements, by comparing (1) the probabilities of bats visiting the focal trees each night immediately after  the manipulation, against (2) the probability of bats visiting the target tree during non-fruit periods.

1. Get the probability of bats tracked during the manipulation to visit the focal trees 1-6 nights after the manipulation:

```{r}

syc_man<-data.frame("fruits"="manipulation" ,"sum_bats_vis"=sum(manip$n_visited_either_tree,na.rm=T)  ,"sum_bats_all"=sum(manip$n_total,na.rm=T))
syc_man$sum_bats_not<-syc_man$sum_bats_all-syc_man$sum_bats_vis

all_manip_tags<-syc_man$sum_bats_all

man_tags_nights<-manip_syc_tags %>% dplyr::group_by(Nights_after_manipulation, manip) %>% summarise("sum_bats_night"=sum(length(unique(tags))))
fifth_night<-data.frame("Nights_after_manipulation"=c(5,5),"manip"= "naive","sum_bats_night"=0)

man_tags_nights<-rbind(man_tags_nights,fifth_night)
man_tags_nights<-man_tags_nights[order(man_tags_nights$Nights_after_manipulation),]
# For all bats (naive and not)
#man_tags_nights$accum_bats_night=cumsum(man_tags_nights$sum_bats_night)

man_tags_nights<-man_tags_nights[man_tags_nights$manip=="Naive",]
man_tags_nights$accum_bats_night=cumsum(man_tags_nights$sum_bats_night)

```

**Add fruit phenology infomration to the dataset of bats visitng the trees**
```{r}
syc_dt<-syc_dt[!is.na(syc_dt$date_global),]
start_date<-min(syc_dt$date_global); end_date<-max(syc_dt$date_global)

syc_dt_all<-syc_dt

# add penology
syc_dt_all$fruits<-NA
for (r in 1:nrow(phen_table)){
rows<-which(syc_dt_all$date_global>=phen_table$Start[r] & syc_dt_all$date_global<=phen_table$End[r])
syc_dt_all$fruits[rows]<-phen_table$stage[r]
}
```


Comparing the probability to visit the focal trees from routine movements, which are calculated by sampling time-intervals, from 1-6, and calculating the proportion of bats visiting the trees based on the summation of the number of bats visiting and not-visiting in multiple time intervals that, together, result in at least 70 bats tracked, the number corresponding directly to the sample size of the manipulation. 

```{r}
non_fruit_p <- na.omit(phen_table[phen_table$stage == "no_fruits", ])
syc_dt_nof <- syc_dt %>% rowwise() %>%
  filter(any(date_global >= non_fruit_p$Start & date_global <= non_fruit_p$End))

test_intervals <- c(1:6)

# Initialize a data frame to store the count of intervals and aggregated intervals
interval_summary <- data.frame(test_interval = integer(),
                               n_intervals = integer(),
                               n_aggregated_intervals = integer(),
                               total_bats = integer())

#ii<-6
total_bats_ts<-70
for (ii in test_intervals) {
    tag_tr <- 2  # minimum number of tags to be tracked in the time-interval
    test_interval <- ii

    all_possible_intervals <- lapply(1:nrow(non_fruit_p), function(i) {
        seq(non_fruit_p$Start[i], non_fruit_p$End[i] - test_interval, by = "day")
    }) %>% unlist() %>% as.Date(origin = "1970-01-01")

    possible_starts <- Filter(function(start_date) {
        end_date <- start_date + test_interval - 1
        interval_data <- syc_dt_nof %>%
          filter(date_global >= start_date & date_global <= end_date)
        
        total_bats <- length(unique(interval_data$TAG))
        all(seq(start_date, end_date, by = "day") %in% syc_dt_nof$date_global) && total_bats >= tag_tr
    }, all_possible_intervals)

    proportions <- vector()  # Initialize empty vector for storing proportions
    remaining_starts <- possible_starts  # Initialize with all possible starts
    n_aggregated_intervals <- 0  # Initialize counter for aggregated intervals

    while (length(remaining_starts) > 0) {
    total_bats = 0
    n_intervals_this_aggregation = 0
    n_visited = 0  # Initialize before the aggregation loop
    n_not_visited = 0  # Initialize before the aggregation loop

        while (total_bats < total_bats_ts && length(remaining_starts) > 0) {
            selected_index <- sample(seq_along(remaining_starts), 1)
            random_date <- remaining_starts[selected_index]
            remaining_starts <- remaining_starts[-selected_index]

            end_date <- random_date + test_interval - 1  

            interval_data <- syc_dt_nof %>%
                filter(date_global >= random_date & date_global <= end_date)

            n_visited_interval <- length(unique(interval_data$TAG[interval_data$syc == "yes"]))
        n_not_visited_interval <- length(unique(interval_data$TAG[interval_data$syc == "no"]))

        n_visited = n_visited + n_visited_interval
        n_not_visited = n_not_visited + n_not_visited_interval
        total_bats = total_bats + n_visited_interval + n_not_visited_interval
        n_intervals_this_aggregation <- n_intervals_this_aggregation + 1
        }

        if (total_bats >= total_bats_ts) {
        proportions = c(proportions, n_visited / total_bats)
        n_aggregated_intervals <- n_aggregated_intervals + n_intervals_this_aggregation
            
            # Record the results for this test interval
    interval_summary <- rbind(interval_summary, data.frame(
        test_interval = test_interval,
        n_intervals = length(possible_starts),
        n_aggregated_intervals = n_aggregated_intervals,
        total_bats = total_bats
    ))
        }
    }

    
 proportions <- na.omit(proportions)
hist( proportions, breaks=25)
 file_name <- paste0("proportions_no_replacement/", "proportion", test_interval, "_Nights_at_least_",total_bats_ts,".rds")
    saveRDS(proportions, file = file_name)
    
    interval_summary
    saveRDS(interval_summary, file = paste0("proportions_no_replacement/interval_summary_",ii,"_Nights.rds"))

}
```

**Plot results of permutations and calculate p-values into a summary table ("results")**
```{r}
# With all bats, including smeared ones
#man_tags_nights$observed_proportion <- man_tags_nights$accum_bats_night /71

# For non-manipulated bats only:
man_tags_nights$observed_proportion <- man_tags_nights$accum_bats_night /56

results <- data.frame()  # Initialize an empty data frame to store results

for (i in 1:6) {
    proportions_file <- sprintf("data_R/proportions_no_replacement/proportion%d_Nights_at_least_70.rds", i)
    proportions <- readRDS(proportions_file)
    
    manipulation_proportion<-man_tags_nights$observed_proportion[man_tags_nights$Nights_after_manipulation==i]
    # Load the corresponding interval summary for this test interval
    summary_file <- sprintf("data_R/proportions_no_replacement/interval_summary_%d_Nights.rds", i)
    interval_summary <- readRDS(summary_file)
    
    # Calculate the mean number of bats per interval and total intervals taken from the summary
    mean_bats_per_interval <- mean(interval_summary$total_bats[interval_summary$test_interval==i])
    n_intervals_taken <- max(interval_summary$n_aggregated_intervals[interval_summary$test_interval==i])

    # Plotting
    ggplot() +
      geom_histogram(aes(x = proportions), binwidth = 0.005, fill = "grey", alpha = 0.7) +
      geom_vline(aes(xintercept = manipulation_proportion), color = "red", linetype = "dashed", size = 1.5) +
      theme_classic() + theme(text = element_text(size=20)) +
      labs(x = "Proportions of Bats Visiting Focal Trees", y = "Frequency")


   
     # Calculating the p-value for a one-tailed test
    p_value <- sum(proportions >= manipulation_proportion) / length(proportions)
    #print(p_value)
    
    # Append the results to the results data frame
    results <- rbind(results, data.frame(
        Test_Interval = i,
        Number_of_Intervals_Taken = n_intervals_taken,
        Mean_Number_of_Bats_per_Interval = mean_bats_per_interval,
        Manipulation_Proportion = manipulation_proportion,
        Mean_Permutations_Proportion = mean(proportions),
        P_Value = p_value
    ))
}

# Save the results to file
#saveRDS(results, file = "proportions_no_replacement/results_summary.rds")
print(results)
```

Add a comparison to the proportion of visits to the focal trees of all bats **that were exposed to the manipulation** 2 weeks before the manipulations.
```{r}
manipulation_dates<-na.omit(unique(manip_syc_tags$date_manipulation))
manipulation_dates <- as.Date(manipulation_dates, format = "%d/%m/%Y")


filtered_syc_dt <- data.frame()

# Loop through each manipulation date
for (manip_date in manipulation_dates) {
    # Calculate the start date for two weeks prior
    start_date <- manip_date - 14  # 14 days for two weeks

    # Filter syc_dt for dates within this range and combine with the previous results
    filtered_syc_dt <- rbind(filtered_syc_dt, 
                             syc_dt[syc_dt$date_global >= start_date & syc_dt$date_global < manip_date, ])
}

prop_vis_before_manip<-round(length(filtered_syc_dt$TAG[filtered_syc_dt$syc=="yes"])/length(filtered_syc_dt$TAG),3)
```

Plot histograms vs. observed (manipulation)
```{r}
library(patchwork)

plot_list <- list()  # Initialize an empty list for plots

for (i in 1:6) {
    proportions_file <- sprintf("data_R/proportions_no_replacement/proportion%d_Nights_at_least_70.rds", i)
    proportions <- readRDS(proportions_file)
    
    manipulation_proportion <- man_tags_nights$observed_proportion[man_tags_nights$Nights_after_manipulation == i]

    # Plotting
    p <- ggplot() +
      geom_histogram(aes(x = proportions), binwidth = 0.01, fill = "grey", alpha = 0.7) +
      geom_vline(aes(xintercept = manipulation_proportion), color = "red", linetype = "dashed", size = 1.5) +
      geom_vline(aes(xintercept = prop_vis_before_manip), color = "black", linetype = "dashed", size = 1.5)+
      theme_classic() +
      theme(text = element_text(size = 14)) +
      labs(x = "Proportions of Bats Visiting Focal Trees", y = "Frequency") +
      ggtitle(sprintf("Night %d", i))+ylim(0,10)+xlab("")+ylab("")+xlim(0,0.25)

    plot_list[[i]] <- ggplotGrob(p)
}

# Combine plots into one layout
# Combine plots into one layout
combined_plot <- wrap_plots(plot_list) + 
                 plot_annotation(
                   caption = "Proportions of Bats Visiting Focal Trees",
                   theme = theme(
                     plot.caption = element_text(hjust = 0.5, size = 14),
                     plot.title.position = "plot"
                   )
                 ) 
print(combined_plot)

```


# Plot means and CI of permutation results against manipulation
```{r}
# Initialize a list to store the proportions data
proportions_list <- list()

# Read the proportions data for each night
for (i in 1:6) {
    file_name <- sprintf("data_R/proportions_no_replacement/proportion%d_Nights_at_least_70.rds", i)
    proportions_list[[i]] <- readRDS(file_name)
}

# Prepare the permutation results data frame with mean and standard deviation
permutation_results <- data.frame(
  Night = 1:6,
  Mean = sapply(proportions_list, mean),
  SD = sapply(proportions_list, sd)
)
# For all bats
#man_tags_nights$observed_proportion <- man_tags_nights$accum_bats_night /71

# For manipulated bats only:
man_tags_nights$observed_proportion <- man_tags_nights$accum_bats_night /56


# Plotting
ggplot() +
  geom_errorbar(data=permutation_results, aes(x=Night, ymin=Mean - SD, ymax=Mean + SD), width=.1) +
  geom_point(data=permutation_results, aes(x=Night, y=Mean), size=4, color="#333340") + geom_line() +
  geom_point(data=man_tags_nights, aes(x=Nights_after_manipulation, y=observed_proportion), size=4, color="#ad2895") + geom_line() +
  labs(x="Number of Nights", y= expression("Proportion Visiting "~ italic("F. sycomorus"))) +
  theme_classic() + theme(text=element_text(size=18))+ ylim(0,0.3) +
  scale_x_continuous(breaks=1:6)


## With Confidence intervals
permutation_results_CI <- tibble(
  Night = 1:6,
  Data = proportions_list
)

permutation_results_CI <- permutation_results_CI %>%
  rowwise() %>%
  mutate(Mean_CI = list(phenesse::mean_ci(Data, bootstraps = 1e+05, conf = 0.95, type = "perc"))) %>%
  unnest(Mean_CI)

# Plotting with confidence intervals
ggplot() +
  geom_errorbar(data=permutation_results_CI, aes(x=Night, ymin=low_ci, ymax=high_ci, color="Routine\nmovement"), width=.1) +
  geom_line(data=permutation_results_CI, aes(x=Night, y=estimate, color="Routine\nmovement"), size=1) +
  geom_point(data=permutation_results_CI, aes(x=Night, y=estimate, color="Routine\nmovement"), size=3) +
  geom_line(data=man_tags_nights, aes(x=Nights_after_manipulation, y=observed_proportion, color="Manipulation"), size=1) +
  geom_point(data=man_tags_nights, aes(x=Nights_after_manipulation, y=observed_proportion, color="Manipulation"), size=3) +
  labs(x="Night", y=expression("Proportion Visiting "~ italic("F. sycomorus"))) +
  scale_color_manual(values=c("Routine\nmovement" = "blue", "Manipulation" = "red")) +
  guides(color=guide_legend(title=NULL)) +  # Remove legend title
  theme_classic() +
  theme(text=element_text(size=18)) +
  ylim(0, 0.3)+
    geom_hline(yintercept=prop_vis_before_manip, linetype="dashed", color="grey", size=1)  


```
